---
title: "Exploring CMG Eye data slice"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Exploring CMG Eye data slice}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction

The CMG Eye workspace was created to work with a small collection
of records from CMG, created through Gen3 operations using
`gen3.theanvil.io/explore`.  See the appendix for some special
considerations about using RStudio in AnVIL for this particular
workbook.

Our first objective is to review tables of subject and sample
characteristics.

```{r start}
suppressPackageStartupMessages({
  library(AnVIL)
  library(DT)
  library(ggplot2)
  })
avworkspace("use-strides/CMG Eye")  # declares until reset
tabs = avtables()
tabs
```

## The subject table

```{r stab}
subjtab = avtable("subject")
dim(subjtab)
names(subjtab)
```

## The sample table

```{r satab}
samptab = avtable("sample")
dim(samptab)
names(samptab)
```

Let's have a searchable table:
```{r dodt}
datatable(samptab)
```

# Working with the sequencing data

## Basic data table

```{r lkseq}
seqtab = avtable("sequencing")
datatable(seqtab)
```

Check the distribution of sequencing depths:
```{r dohist}
ggplot(seqtab, aes(x=`pfb:total_reads`)) + geom_histogram()
```

## Acquire a CRAM file and its index

Here we pick records using an identifier prefix for a single contributor,
and convert DRS to google storage identifiers.
```{r filt}
drs_pair = seqtab %>% filter(grepl("PIE_OGI001_000346_1", `pfb:seq_filename`)) %>%
    select(`pfb:ga4gh_drs_uri`) %>% as.data.frame
drs_pair[[1]]
gsu = drs_stat(drs_pair[[1]])$gsUri
gsu
```
We'll use AnVIL to copy the CRAM and CRAI to our local folder.
```{r drscp}
drs_cp(drs_pair[[1]], ".")
dir(patt="cram")
```

## Filter a selection of CRAM files based on phenotype

```{r dojoin}
mm = left_join(subjtab, mutate(seqtab, 
  `pfb:participant_id`=gsub(".cram", "", `pfb:file_name`)), 
  by="pfb:participant_id")
dim(mm)
mm %>% select(`pfb:phenotype_description`, `pfb:seq_filename`, `pfb:ga4gh_drs_uri`) %>%    
  filter(`pfb:phenotype_description`=="Retinitis pigmentosa")
```

# Conclusions

This document illustrated simple activities with AnVIL package and AnVIL
data from CMG.




# Appendix

This workbook is generated within RStudio in AnVIL as a project that is version-controlled.
The repo `github.com/vjcitn/CMG_EYE` is used to manage the R markdown.  Every change executed
in the RStudio editor
can be committed and pushed to the repo.  Changes made elsewhere can be pulled into the project via git.
It was necessary to use a personal access token to control the access.